---
title: "CovidAnalysis"
author: "Shamalan"
date: "11/29/2020"
output:
  pdf_document: default
  html_document: default
---

## R Markdown

```{r}
library(dplyr)
library(ggplot2)
library(ggraph)
library(tidyverse)  # data manipulation and visualization
library(modelr)     # provides easy pipeline modeling functions
library(broom)      # helps to tidy up model outputs
library(tidyr)
```
##Reading datasets


#https://github.com/OxCGRT/covid-policy-tracker/blob/master/documentation/codebook.md.
```{r}
ox <- read.csv(file.choose())
```

#https://www.google.com/covid19/mobility/
```{r}
google <- read.csv(file.choose())

```

#https://ourworldindata.org/coronavirus/country/malaysia?country=~MYS
```{r}
owid <- read.csv(file.choose())
```

#the link connected to drive because the dataset is cleaned and merged using another tool
#https://drive.google.com/file/d/171iSft5sjYHhDANZklI02i35dM8U1bSe/view?usp=sharing
```{r}
google_ox <- read.csv(file.choose())
```

## Data Cleaning

#ox, google and owid dataset cleaning
```{r}
ox = subset(ox, select=-c(RegionName, RegionCode, M1_Wildcard))
ox<- filter(ox, CountryName == "Malaysia")

google = subset(google, select=-c(sub_region_2, metro_area, census_fips_code,iso_3166_2_code))

owid = subset(owid, select=-c(iso_code, continent, new_cases_smoothed, new_deaths_smoothed ,total_cases_per_million, new_cases_per_million, new_cases_smoothed_per_million,
                              total_deaths_per_million, new_deaths_per_million, new_deaths_smoothed_per_million, icu_patients_per_million, hosp_patients_per_million,
                              weekly_icu_admissions_per_million, weekly_hosp_admissions_per_million, total_tests_per_thousand, new_tests_per_thousand, new_tests_smoothed,
                              new_tests_smoothed_per_thousand))
owid<- filter(owid, location == "Malaysia")
```

#combine google and owid dataset then clean it
```{r}
google_owid <- merge(owid, google, by='date')
str(google_owid)
google_owid = subset(google_owid, select=-c(reproduction_rate, icu_patients, hosp_patients, weekly_icu_admissions, weekly_hosp_admissions, total_tests, new_tests, tests_per_case,
                                            positive_rate, tests_units, stringency_index, population, population_density, median_age, aged_65_older, aged_70_older, gdp_per_capita,
                                            extreme_poverty, cardiovasc_death_rate, diabetes_prevalence, female_smokers, male_smokers, handwashing_facilities, hospital_beds_per_thousand,
                                            life_expectancy, human_development_index))
```

#Descriptive Analysis

#Plotting of community mobility across the dates

```{r}
filtered <- filter(google, sub_region_1 == "")
filtered$date <- as.Date( filtered$date, '%m/%d/%Y')
require(ggplot2)

library("tidyverse")
filtered_df <- filtered %>%
  select(date, grocery_and_pharmacy_percent_change_from_baseline, retail_and_recreation_percent_change_from_baseline, parks_percent_change_from_baseline, 
         transit_stations_percent_change_from_baseline, workplaces_percent_change_from_baseline, residential_percent_change_from_baseline) %>%
  gather(key = "variable", value = "value", -date)

ggplot(filtered_df, aes(x = date, y = value)) + 
  geom_line(aes(color = variable)) + 
  scale_color_manual(values = c("darkred", "orange", "black", "darkgreen", "steelblue", "purple"))
```

#We used the ggplot function to plot the percentage change of mobility of Malaysian community in different places. The rationale behind this was to compare and 
#see the overall trend of the mobility and search of insights or the behaviour of the community.
#The generated plot showed the distribution of Malaysian community mobility across the date. We can observe that somewhere near March the community’s 
#mobility had a huge downturn for all places categories except residential. This phenomena was due to the implementation of Movement Control Order(MCO) 
#by the Malaysian government which restricts the movement of the community from visiting all places except groceries and pharmacies. After the huge downturn, 
#we can see that the trend maintained for a while and started to rise back at around May. This is because the number of new covid cases in Malaysia had decreased 
#and was under control and thus the government lifted MCO and implemented Conditional Movement Control Order (CMCO) which allows most shops and business 
#to continue operation with some rules for the community to follow with. Then, the next obvious trend we are able to capture is another downturn of all 
#places except residential once again happening at 1st of October. This is explained by the huge number of cases recorded in Sabah and soon the implementation 
#of a new restrictions control soon at 14th of October whereby the downturn had a dip.

#New cases across the date
```{r}
filtered2 <- owid
filtered2$date <- as.Date(filtered2$date, '%m/%d/%Y')
ggplot( data = filtered2, aes( date,new_cases )) + geom_line() 
```


#GGplot is used for this plot in hopes of identifying the new cases recorded in each day and identify why and how it happened
#The generated plot showed the number of new cases across the date. In somewhere March, we saw a spike in which that was the first wave of the pandemic 
#in Malaysia. Then somewhere in May, we see the plot had a downturn which indicates the number of new cases is decreasing. Then, the trend maintained for 
#a few more months until September. We can see that the trend starts to rise again in September where there is a outbreak occurring in Sabah after the election which is 
#identified as the 3rd wave of the pandemic in Malaysia. Soon enough reaching October, the number of new cases had a huge spike and was lost of control 
#whereby almost everyday after October is breaking the record most cases recorded per day reaching 2000 cases.

#####################################

#Predictive Analysis

#Predictive analysis is a branch under advanced analytics whereby the main purpose is to make predictions of unknown events occurring in the future. 
#Predictive analysis may use many techniques to analyse the current data for future predictive and the method we used here is modelling. 
#There are 2 future or objectives we are trying to predict based on our dataset which were : 
#1.	The lifestyle of Malaysians in the future as number of new cases growing
#2.	The effect of government implemented policies against Malaysians lifestyle during the pandemic.

# 1.	The lifestyle of Malaysians in the future as number of new 

```{r}

sample <- sample(c(TRUE, FALSE), nrow(google_owid), replace = T, prob = c(0.6,0.4))
train <- google_owid[sample, ]
test <- google_owid[!sample, ]

model1 <- lm( new_cases ~ retail_and_recreation_percent_change_from_baseline + grocery_and_pharmacy_percent_change_from_baseline +
                parks_percent_change_from_baseline + transit_stations_percent_change_from_baseline + workplaces_percent_change_from_baseline +
                residential_percent_change_from_baseline, data=train)
summary(model1)
tidy(model1)
```


#For research question 1, we merged the “region mobility report“ dataset with “owid-covid-data” dataset to obtain the percentage change of Malaysian lifestyle 
#and the number of new cases.  While for research question 2, we merged the  “region mobility report” with “OxCGRT” dataset to obtain government policies and 
#the lifestyle of Malaysians.

#We used the “tidyverse”, “modelr”, and “broom” libraries for data manipulation, easy pipeline modelling function, and mode output tidying.  Then we split the 
#data to 60 % for training and 40% for validation and built a linear regression model with new cases as response variable with 
#“retail_and_recreation_percent_change_from_baseline” ,  “grocery_and_pharmacy_percent_change_from_baseline”, “parks_percent_change_from_baseline”,
#“transit_stations_percent_change_from_baseline”, “residential_percent_change_from_baseline” and “workplaces_percent_change_from_baseline” as predictor variables 
#to fit the research question. The output of the model shows that grocery and pharmacy percent change is insignificant to the contribution of new cases since the 
#p-value is 0.058651 which is above 0.5. We could interpret this as grocery and pharmaceutical purchasing was done by 1 individual in a whole big family which 
#unlike recreation, parks, workplaces require the participation of each individual. However, if we further investigate this model, it reported a adjusted r-squared 
#value of 0.1264 which says that this model does not fit close to the regression line and proved that it is a weak model to predict the number of new cases based 
#on the rate of Malaysians moving around during the pandemic. However, if we look back to the research question, the main objective is to investigate the lifestyle 
#of Malaysians in the future as number of new cases grows. Although we cant get to generate a model to prove that new cases is highly related to lifestyle and 
#predict the future lifestyle since it is a weak model, we are able to capture an interesting phenomena which is – Malaysians would rather continue their previous 
#lifestyle as before the pandemic than following the best practice of staying at home to prevent own self from getting the virus. We are used to our normal 
#lifestyle where we had the freedom to do as what we want which fast forward to today this freedom was encaged with many rules and regulation. Thus, we predict 
#that in the future if there was to be another pandemic to occur, the Malaysian community would not be cautious and obeying the rules strictly as they had a 
#previous experience of combating the virus and live within its presence.



#2.	The effect of government implemented policies against Malaysians lifestyle during the pandemic.
#For research question 2, we are interested to find the effect of government implemented policies against Malaysians lifestyle during the pandemic and 
#predict the future based on it. Thus, again we built a regression model using government policies as predictor variable and community mobility in each 
#sector as response. The first model was built using retail and recreation percentage change as response variable


#cleaning of dataset
```{r}
google_ox = subset(google_ox, select=-c(CountryName, CountryCode, C1_Flag, C2_Flag ,C3_Flag, C4_Flag, C5_Flag,
                                        C6_Flag, C7_Flag, E1_Flag, H1_Flag, H6_Flag, M1_Wildcard, ConfirmedCases, ConfirmedDeaths, StringencyIndex, StringencyIndexForDisplay
                                        , StringencyLegacyIndex, StringencyLegacyIndexForDisplay, GovernmentResponseIndex, GovernmentResponseIndexForDispla, ContainmentHealthIndex, 
                                        ContainmentHealthIndexForDisplay, EconomicSupportIndex, EconomicSupportIndexForDisplay))

sample2 <- sample(c(TRUE, FALSE), nrow(google_ox), replace = T, prob = c(0.6,0.4))
train2 <- google_ox[sample2, ]
test2 <- google_ox[!sample2, ]
```

#regression model of predicting the RETAIL AND RECREATION visits while government policy implementation
```{r}
model3 <- lm(retail_and_recreation ~ C1_School.closing + C2_Workplace.closing + C3_Cancel.public.events + C4_Restrictions.on.gatherings + C5_Close.public.transport +
               C6_Stay.at.home.requirements + C7_Restrictions.on.internal.move + C8_International.travel.controls , data=train2)
summary(model3)
tidy(model3)

```
#The result showed that “C3 Cancel public events” and “C7 Restrictions on internal move” had a p-value of greater than the significant figure of 0.05 at 0.5249 
#and 0.3836 respectively while “C5_Close.public.transport” had did not provide any information because all the values in the dataset is 0. Thus, these 3 variables 
#were excluded from the model and a new regression model was built.


```{r}
model3.2 <- lm(retail_and_recreation ~ C1_School.closing + C2_Workplace.closing + C4_Restrictions.on.gatherings +
                 C6_Stay.at.home.requirements + C8_International.travel.controls , data=train2)
summary(model3.2)
tidy(model3.2)
```

#This time, the remaining variables in the model were “C1 School closing”, “C2 Workplace closing”, “C4 Restrictions on gathering”, “C6 Stay at home requirements”,and “C8 International travel controls”. All of them pad a p-value of lesser than 0.05 which proved that they are significant in predicting the number of new cases. Alongside this model, we had an F value of 95.21 and adjusted R-squared value of 0.7536. Thus, we are able to say that this is a strong model in predicting the new cases with 5 variables(policies) having a significant impact on it. 

#Based on the generated output, we managed to interpret several insights on it. It is seen that the closure of school, workplace, restrictions on gathering, stay at home requirements, international travel controls is working well in decreasing Malaysians’ lifestyle of visiting to retails and recreation sites. This is seen and proved working in the first wave of the pandemic whereby the government shut down schools, workplace, restrict movements, cancel international flights managed to decrease retail and recreation visits as well as reducing the number of new cases to single digits. However, as the 3rd wave starts to occur, the government did not take any heavy measures in controlling the community like forcefully Work-From-Home, closure of schools immediately, business restrictions. As an effect, the community continued to their usual visits to retail and recreation sites while the daily new cases had an upturn and finally under control at 1000+- cases daily. This is a very good evidence in how powerful is government policy implementation is to controlling the community’s lifestyle directly and number of new cases indirectly. In the near future, if we wish to bring down the number of daily new cases, the government will need to stricten back the rules and regulations as well as containment and closure policies. That way, the people can only have less activity in their mobility and only the lesser contacts will occur and new cases will see a downturn. Of course the implementation of strict rules and closure policies will affect the operation of a myriad of business 
#in Malaysia which in turn provides a big hit to the overall economy, but it all depends on the government’s priority. If the priority is to decrease the 
#number of cases, strict rules must be applied and the overall economy had to be sacrificed. In the other hand, if the overall economy is still the priority,
#then the community can only hope for the best in reducing close contacts with others while protecting own self while getting used to this new norm of living. 



#Diagnostic Analysis#############
#Diagnostic Analysis utilises the data from the descriptive analysis for further insights. Throughout this analysis, the focus would be to understand the data in detail. The information obtained is to best understand the covid pandemic statistics in Malaysia as well as to draw comparison between Malaysia and the nearby ASEAN countries.

#once open the link, please right click and save it as Microsoft Excel Comma Seperated Values File
#https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv
```{r}
confirmedCovid <- read.csv(file.choose())
str(confirmedCovid)
```
#once open the link, please right click and save it as Microsoft Excel Comma Seperated Values File
#https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv
```{r}
deathsCovid <- read.csv(file.choose())
```

#once open the link, please right click and save it as Microsoft Excel Comma Seperated Values File
#https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv
```{r}
recoveredCovid <- read.csv(file.choose())
```


# DATA CLEANING

```{r}
# Convert each data set and aggregate at country level

confirmedCases <- confirmedCovid %>% gather(key="date", value="confirmed", -c(Country.Region, Province.State, Lat, Long)) %>% group_by(Country.Region, date) %>% summarize(confirmed=sum(confirmed))
deathsCases <- deathsCovid %>% gather(key="date", value="deaths", -c(Country.Region, Province.State, Lat, Long)) %>% group_by(Country.Region, date) %>% summarize(deaths=sum(deaths))
recoveredCases <- recoveredCovid %>% gather(key="date", value="recovered", -c(Country.Region, Province.State, Lat, Long)) %>% group_by(Country.Region, date) %>% summarize(recovered=sum(recovered))
summary(confirmedCases)

# Combining all 3 datasets
masterDataset <- full_join(confirmedCases, deathsCases) %>% full_join(recoveredCases)

# Covert the character variable to date
str(masterDataset) 
masterDataset$date <- masterDataset$date %>% sub("X", "", .) %>% as.Date("%m.%d.%y")
str(masterDataset) 
# Identifying the number of days
masterDataset <- masterDataset %>% group_by(Country.Region) %>% mutate(cumconfirmed=cumsum(confirmed), days = date - first(date) + 1)

```

#Extracting the Worldwide and Malaysian data

```{r}
worldwide <- masterDataset %>% group_by(date) %>% summarize(confirmed=sum(confirmed), cumconfirmed=sum(cumconfirmed), deaths=sum(deaths), recovered=sum(recovered)) %>% mutate(days = date - first(date) + 1)

msia <- masterDataset %>% filter(Country.Region=="Malaysia")

```

#Summary Statistics of the masterDataset

```{r}
summary(masterDataset)
by(masterDataset$confirmed, masterDataset$Country.Region, summary)
by(masterDataset$cumconfirmed, masterDataset$Country.Region, summary)
by(masterDataset$deaths, masterDataset$Country.Region, summary)
by(masterDataset$recovered, masterDataset$Country.Region, summary)
summary(msia)
```

#Line graph with a log10 scale
#Logarithmic scales are useful for visualizing data when the difference between measures is comparatively large.
```{r}
ggplot(msia, aes(x=days, y=confirmed)) + geom_line() +
  theme_classic() +
  labs(title = "Covid-19 Malaysia Confirmed Cases", x= "Days", y= "Daily confirmed cases  (log scale)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(trans="log10")
```

#The line graph gives a general idea of the number of cases as the days pass. The y-axis uses the Log10 scale as it would be best to respond to the skewness towards large values. This graph would laso allow for the percent change and multiplicative factors to be easily read. The log10 scale is also used because we are dealing with data of large values.
#The line graoh indicates that there is a steep increase from day 0 to 100 followed by a steady increase from day 100 to 250 and finally a significant increase towards 300 days.

#Comparison graph between selected ASEAN countries
```{r}
countryselection <- masterDataset %>% filter(Country.Region==c("Malaysia", "Singapore", "Indonesia", "Thailand", "Cambodia"))
ggplot(countryselection, aes(x=days, y=confirmed, colour=Country.Region)) + geom_line(size=1) +
  theme_classic() +
  labs(title = "Covid-19 cases of nearby countries", x= "Days", y= "Daily confirmed cases (log scale)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(trans="log10")
```

#This graph compares the Covid cases between Malaysia and its nearby countries (Cambodia, Indonesia, Singapore and Thailand). This graph indicates that Malaysia has gotten a moderate number of cases when compared to the other countries. However, as noted earlier, there is a sharp increase towards day 300. This may be caused by many different reasons. Malaysia's cases has taken a sharp increase and equaled the cases that of Singapore.

#Detailed comparison between the confirmed, unconfirmed, deaths and recovered amongst the chosen ASEAN countries.
```{r}
str(countryselection)
countryselection %>% gather("Type", "Cases", -c(date, days, Country.Region)) %>%
  ggplot(aes(x=days, y=Cases, colour=Country.Region)) + geom_line(size=1) +
  theme_classic() +
  labs(title = "Covid-19 Cases by Country", x= "Days", y= "Daily cases (log scale)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(trans="log10") +
  facet_grid(rows=vars(Type))
```

#This matrix graph gives a brief comparison between the confirmed, cumulative confirmed, deaths and the recovered cases. This graph shows Malaysia's performance when it comes to its comparison to the various ASEAN countries. 
#The cumulative confirmed cases indicates the total number of Covid cases based on the days. When it comes to the number of deaths, Malaysia is seen to have the second highest number of deaths compared to the chosen ASEAN countries. 
#The number of recovered however, places Malaysia at a steady position on top compared to the chosen ASEAN countries.

